Concurrent Programming with CXL

### 2020-05-17 book/paper.tex
Added text about progress property

\begin{code}
\begin{verbatim}
while True:
    # enter critical section
    @cs: pass;  # in critical section
    # exit critical section
;
\end{verbatim}
\end{code}

Here \texttt{@cs} is a \emph{label}, identifying a location in the
CXL machine code.  The first thing we need to ensure is that there
cannot be two processes at the critical section.  We would like to
place an assertion at the \texttt{@cs} label that specifies that
only the current process can be there.  CXL in fact supports this.
It has a method \texttt{atLabel($L$)}, where $L$ is the string name
of a label (in this case, \texttt{"cs"}).  The method returns a
bag of name tags of processes executing at that label.  The bag is
represented by a dictionary that maps each element in the bag to the
number of times the element appears in the bag.

@@

>>>>

\begin{code}
\begin{verbatim}
while True:
    @ncs: pass; # in non-critical section
    # enter critical section
    @cs: pass;  # in critical section
    # exit critical section
;
\end{verbatim}
\end{code}

Here \texttt{@ncs} and \texttt{@cs} are \emph{labels}, each identifying
a location in the CXL machine code.  The first thing we need to
ensure is that there cannot be two processes at the critical section.
We would like to place an assertion at the \texttt{@cs} label that
specifies that only the current process can be there.  CXL in fact
supports this.  It has a method \texttt{atLabel($L$)}, where $L$
is the string name of a label (in this case, \texttt{"cs"}).  The
method returns a bag of name tags of processes executing at that
label.  The bag is represented by a dictionary that maps each element
in the bag to the number of times the element appears in the bag.

@@

However, mutual exclusion by itself easy to ensure.
For example, we could insert the following code to enter the
critical section:
\begin{code}
\begin{verbatim}
while True: pass;;
\end{verbatim}
\end{code}
This code will surely prevent two or more processes from being
at label \texttt{cs} at the same time.
But it does so by preventing \emph{any} process from reaching
the critical section.
The point is that we need another property besides mutual exclusion.

Mutual exclusion is an example of a \emph{safety property}, a
property that ensures that \emph{nothing bad will happen}, in this case
two processes being in the critical section. Now
we need a \emph{liveness property}: we want to ensure that
\emph{eventually something good will happen}.
There are various way of specifying this,
but here we will propose the following informally: when nobody is in the
critical section, then any process should be able to move forward and
enter the critical section.  We call this \emph{progress}.
